# Reproducible Research: Course Project 1

##Loading and preprocessing the data

###Code for reading in the dataset and/or processing the data
```{r readfile, echo=TRUE}
step1 <- read.table("C:/Users/ndechativong/Documents/R/coursera/data/activity.csv", header = TRUE, sep = ",")
head(step1)
```

##What is mean total number of steps taken per day?

###Histogram of the total number of steps taken each day
```{r hist_stepday, echo=TRUE}
step1_sum <- aggregate(step1$steps, by=list(date=step1$date), sum, na.rm = TRUE)
hist(step1_sum$x, main = "Histogram of the total number of steps taken each day", xlab = "Number of steps")
```

###Mean and median number of steps taken each day
```{r mean_median_stepday, echo=TRUE}
step1_mean <- mean(step1_sum$x, na.rm = TRUE)
step1_mean

step1_median <- median(step1_sum$x, na.rm = TRUE)
step1_median
```

##What is the average daily activity pattern?

###Time series plot of the average number of steps taken
```{r timeplot_avg, echo=TRUE}
library("lattice")
step1_int_avg <- aggregate(step1$steps, by=list(interval=step1$interval), mean, na.rm = TRUE)
xyplot(step1_int_avg$x ~ step1_int_avg$interval, type = "l", xlab = "5-minute interval", ylab = "Average number of steps")
```


###The 5-minute interval that, on average, contains the maximum number of steps
```{r max_interval, echo=TRUE}
apply(step1_int_avg, MARGIN = 2, function(x) max(x, na.rm = TRUE))
```

##Imputing missing values

###Calculate and report the total number of missing values in the dataset
```{r sum_missinputs, echo=TRUE}
sum(is.na(step1$steps))
```

###Code to describe and show a strategy for inputing missing data
```{r input_miss, echo=TRUE}
# Average steps per day
step1_day_avg <- aggregate(step1$steps, by=list(date=step1$date), mean, na.rm = TRUE)

# Number of missing inputs per day
aggregate(is.na(step1$steps), by=list(date=step1$date), sum, na.rm = FALSE)

# Create new data frame to capture made-up input values
step2 <- step1

for (n in 1:nrow(step1)) {
    s <- step1[n, ]
    
    if (is.na(s$steps))
        steps <- subset(step1_int_avg, interval == s$interval)$x
    else
        steps <- s$steps

    step2[n, 1] <- round(steps)
}
```

###Histogram of the total number of steps taken each day after missing values are inputed
```{r hist_summiss, echo=TRUE}
step2_sum <- aggregate(step2$steps, by=list(date=step2$date), sum, na.rm = TRUE)
hist(step2_sum$x, main = "Histogram of the total number of steps taken each day", xlab = "Number of steps")
```

###Mean and median number of steps taken each day after missing values are inputed
```{r mean_miss, echo=TRUE}
step2_mean <- mean(step2_sum$x, na.rm = TRUE)
step2_mean

step2_median <- median(step2_sum$x, na.rm = TRUE)
step2_median
```

##Are there differences in activity patterns between weekdays and weekends?


###Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r plot_avg_week, echo=TRUE}
step2$weekday <- weekdays(as.Date(step2$date))
step2$daytype <- ifelse((step2$weekday == "Saturday" | step2$weekday == "Sunday"), "Weekend", "Weekday")
step2_int_avg <- aggregate(step2$steps, by=list(interval=step2$interval, step2$daytype), mean, na.rm = TRUE)
names(step2_int_avg) <- c("interval", "daytype", "step")

xyplot(step ~ interval | daytype, step2_int_avg, type = "l", layout = c(1, 2), xlab = "5-minute interval", ylab = "Average number of steps")
```



